---
title: "425 Final Project"
date: "December 13, 2017"
output: html_document
---
Team Members : <br>
1. Ankush Agrawal (ankusha2) <br>
2. Apurv Garg (apurvg2) <br>
3. Xiao Liu (xiaol5) <br>

###About the data
The dataset used for analysis was sourced from Airbnb listings for Chicago from Inside Airbnb [link](http://insideairbnb.com/get-the-data.html). <br>
The data consists of 95 variables and 5207 observations where each observation represents a listing.

The data is available at the Inside Airbnb site and is scrapped from publicly available information from the Airbnb site. We are using the detailed listings data for Chicago. It was scrapped on 10^th^ May 2017.

In this project, we have done the following :
<br>1. Pre-process the data for the purpose of EDA and modelling. We have treated the data for missing values, outliers, erroneous data (values which do not make sense). This will also help us understand important features of the data like the price of a listing, number of reviews, reviews per month(acts as proxy for booking rate), etc.
<br>2. Use statistical modelling  techniques to find interesting insights from the data.
<br>3. Create an interactive dashboard so that the end user can get insights that we have collected in the form of charts, tables and maps.

###Getting Started : Data Preprocessing

```{r}
Chicago <- read.csv("listings 2.csv", header = TRUE,stringsAsFactors = F)

a <- strsplit(Chicago$amenities,",")
```

Looking at the summary statistics of the table, we observe the following things : <br>
a. A few variables like scrape_id are either entirely null or have only one single value populated throughout (e.g. experiences_offered) and hence, it can be removed.<br>

```{r}
# Removing variables that contain only a single value
Chicago <- Chicago
a <- apply(Chicago,2,unique)
b <- NULL
for(i in 1:length(a))
{
  if(length(a[[i]])==1){
    b <- rbind(b,names(a[i]))
  }
}

b <- as.vector(b)
Chicago[,b] <- NULL
```

This leads to a removal of 9 such variables. <br>

b. All the url related variables can be removed as they contain weblinks which aren't useful for any kind of analysis<br>
```{r}
Chicago[,names(Chicago)[grep("url",names(Chicago))]] <- NULL
```
This leads to removal of 8 such variables. <br>

c. Next we remove all those variables which have more than 80% missing values<br>
```{r}
a <- apply(Chicago,2,is.na)
b <- apply(a,2,sum)
Chicago[,names(b[b/5207>=0.80])] <- NULL
```

This leads to removal of 1 variable.
<br>
d. Based on intuition and further inspection we observed that variables like security_deposit, weekly_price, monthly_price, first_review, last_review,  jurisdiction_names, zipcode, street, market, cleaning_fee,  transit, amenities, house_rules, host_about, neighborhood_overview, host_neighbourhood, name, interaction, access, space, notes, summary, description, host_name, host_has_profile_pic, host_verifications, host_neighborhood, require_guest_profile_picture, require_guest_phone_verification, calculated_host_listings_count, host_location, transit,neighborhood_overview,house_rules,host_about,license and requires_license do not add any value to our analysis and should be removed. Hence, we have removed 37 such variables.<br>

```{r}
Chicago[,c("security_deposit","weekly_price","monthly_price","first_review","last_review","jurisdiction_names","zipcode","street","market","cleaning_fee","name","interaction","access","space","notes","summary","description","host_name","host_has_profile_pic","host_verifications","host_neighborhood","require_guest_profile_picture","require_guest_phone_verification","calculated_host_listings_count", "host_location","transit","neighborhood_overview","house_rules","host_about","license", "requires_license","host_neighbourhood")] <- NULL
```

e. Removing variables with duplicated values <br>
```{r}
Chicago[,names(Chicago[(duplicated(t(Chicago)))])] <- NULL
```

<br>f. We observe that different listings have different amenities and hence we will create a flag/dummy for the most important amenities we think a user looks for while booking at AirBNB.<br>
These amenities are TV, Internet, Air Conditioning, Breakfast, Kitchen and Pets<br>

```{r,message=FALSE}
# adding amenities categories
a <- Chicago$amenities
Chicago$TV <- ifelse(grepl("TV",a, ignore.case = T)==T,1,0)
Chicago$Internet <- ifelse(grepl("Internet",a, ignore.case= T)==T,1,0)
Chicago$AirCondition <- ifelse(grepl("conditioning",a, ignore.case =T)==T,1,0)
Chicago$Pets <- ifelse(grepl("Pet",a, ignore.case = T)==T,1,0)
Chicago$Pets <- ifelse(grepl("Dog",a, ignore.case = T)==T,1,Chicago$Pets)
Chicago$Pets <- ifelse(grepl("Cat",a, ignore.case = T)==T,1,Chicago$Pets)
Chicago$Kitchen <- ifelse(grepl("Kitchen",a, ignore.case = T)==T,1,0)
Chicago$Breakfast <- ifelse(grepl("breakfast",a, ignore.case = T)==T,1,0)
Chicago[,c("amenities")] <- NULL
```

<br>The data type of the variables which we think might be of significance are converted to the appropriate data types<br>

```{r,message=FALSE,warning=FALSE}
# Converting price variable to integer
Chicago$price <- sub("\\$","",Chicago$price)
Chicago$price <- sub(",","",Chicago$price)
Chicago$price <- as.integer(Chicago$price)

# Changing the default character data type of categorical variable to factor
Chicago$host_response_time <- as.factor(Chicago$host_response_time)
Chicago$host_is_superhost <- as.factor(Chicago$host_is_superhost)
Chicago$host_identity_verified <- as.factor(Chicago$host_identity_verified)
Chicago$neighbourhood_cleansed <- as.factor(Chicago$neighbourhood_cleansed)
Chicago$is_location_exact <- as.factor(Chicago$is_location_exact)
Chicago$property_type <- as.factor(Chicago$property_type)
Chicago$room_type <- as.factor(Chicago$room_type)
Chicago$bed_type <- as.factor(Chicago$bed_type)
Chicago$calendar_updated <- as.factor(Chicago$calendar_updated)
Chicago$instant_bookable <- as.factor(Chicago$instant_bookable)
Chicago$cancellation_policy <- as.factor(Chicago$cancellation_policy)

# Treating host_response_rate and extra_people from character to a numeric variable
Chicago$host_response_rate<- as.numeric(sub("%", "", Chicago$host_response_rate))
Chicago$host_response_rate <- Chicago$host_response_rate/100
Chicago$extra_people <- as.numeric(sub("\\$","",Chicago$extra_people))
```

<br>We see that some listings have 0 as their price, which is not possible and is possibly an erroneous data. Hence, we have removed such listings.<br>

```{r}
# Removing listings which have price = 0
Chicago <- Chicago[-c(717,1334,3093),]
```

<br>We also created a variable with the number of days since the listing is on AirBNB. We leveraged the variable host_since which gives us the data since the listing has been on AirBNB and we subtracted this date from the 10<th> May 2017, i.e. the date when the data was scrapped<br>

```{r}
Chicago$host_since <- as.Date(Chicago$host_since)
Chicago$host_since <- as.Date("2017-05-10")-Chicago$host_since
```

<br>In order to reduce the number of levels in factor variables with large number of levels, we clubbed the factors with low number of listings. In the dataset, we observe that neighbourhood_cleansed, which gives us an idea of the nighbourhood of the listing, has 72 levels to it. 
```{r,message=FALSE}
library(dplyr)
a <- Chicago %>% group_by(neighbourhood_cleansed) %>% summarise(len = length(neighbourhood_cleansed))
#sum(a[(a$len<70),2])

Chicago <- merge(Chicago,a,by = "neighbourhood_cleansed")
Chicago$neighbourhood_cleansed <- as.character(Chicago$neighbourhood_cleansed)
Chicago$neighbourhood_cleansed <- ifelse(Chicago$len<150,"Others",Chicago$neighbourhood_cleansed)
Chicago$len <- NULL
```

###Missing Value Treatment
We checked for all types of missing values in all variables of the final dataset. This includes is.na, "" and "N/A".

```{r}
# checking missing values in different columns
missing = data.frame(col=colnames(Chicago), type = sapply(Chicago, class), 
                     missing = sapply(Chicago, function(x) sum(is.na(x)| x=="" | x=="N/A"))/nrow(Chicago) * 100)

```

We treated the missing values as follows :
```{r}
# Replace missing values in host_response_time with the most common occurring category (its mode)
Chicago$host_response_time<- sub("N/A","within an hour", Chicago$host_response_time)

# replace missing values in host_response rate with mean values observed for host_response_time when it is within an hour
Chicago$host_response_rate <- ifelse(is.na(Chicago$host_response_rate)==T,0.99,Chicago$host_response_rate)

# Replace missing values in bathrooms, bedrooms and beds with the median value
Chicago$bathrooms <- ifelse(is.na(Chicago$bathrooms)==T, 1, Chicago$bathrooms)
Chicago$bedrooms <- ifelse(is.na(Chicago$bedrooms)==T,1,Chicago$bedrooms)
Chicago$beds <- ifelse(is.na(Chicago$beds)==T,1,Chicago$beds)

# Replace the missing values in review scores with the median values as well
Chicago$review_scores_rating <- ifelse(is.na(Chicago$review_scores_rating)==T,97,Chicago$review_scores_rating)
Chicago$review_scores_accuracy <- ifelse(is.na(Chicago$review_scores_accuracy)==T,10,Chicago$review_scores_accuracy)
Chicago$review_scores_cleanliness <- ifelse(is.na(Chicago$review_scores_cleanliness)==T,10,Chicago$review_scores_cleanliness)
Chicago$review_scores_checkin <- ifelse(is.na(Chicago$review_scores_checkin)==T,10,Chicago$review_scores_checkin)
Chicago$review_scores_communication <- ifelse(is.na(Chicago$review_scores_communication)==T,10,Chicago$review_scores_communication)
Chicago$review_scores_location <- ifelse(is.na(Chicago$review_scores_location)==T,10,Chicago$review_scores_location)
Chicago$review_scores_value <- ifelse(is.na(Chicago$review_scores_value)==T,10,Chicago$review_scores_value)
Chicago$reviews_per_month <- ifelse(is.na(Chicago$reviews_per_month)==T,1.55,Chicago$reviews_per_month)

```

###Outliers

```{r,message=FALSE,warning=FALSE}
# Histogram of price
library(ggplot2)
ggplot(data=Chicago, aes(price)) + 
  geom_histogram(fill="red") + 
  labs(title="Histogram of Price") +
  labs(x="Price", y="Count")

# Percentile of price
quantile(Chicago$price, c(.9, .95, .97, 0.975, 0.98, 0.99, 0.995, 0.999, 0.9999))
```


<br>As we can see from the histogram as well as the percentile distribution of Price, there are extreme values in Price. So we performed Winsorization at 99% level and captured the maximum value of price at 650 USD.<br>

```{r}
# Capture the extreme values of Price at 99 percentile level
Chicago$price <- ifelse(Chicago$price>650,650,Chicago$price)
```


###Dashboard
[AirBnB Analysis Dashboard](http://apurvg2.shinyapps.io/airbnb/)<br>

<br>Our UI should contain the following components:<br> 

Map allowing the user to look at the AirBnB listing on a map. There are also filters based on the Price range and Neighbourhoods.<br>
Map on the 2nd tab shows the distribution of listings based on the prices for different room types.<br>
Graphs provide the functionality for the user to have a look at box-plot for different prices and #booking values based on different categorical variables like room_type, bed_type, cancellation policy and propert type.<br>
User can also look at the variation of prices and #bookings vs host since(a scatter plot). We have provided multiple filters like number of bathroom, bedrooms and guests so that user can see the metrics based on their requirements.<br>

**Note: We have used plotly to make graphs which is an interactive graph library. It allows user to subselect factors in the graph by clicking on the legends**<br>

###Analysis 1: Comparing the booking rate based on cancellation policy
<br>We wish to know whether the booking rate is different for different cancellation policies.<br>

H0: The average booking rate changes with the cancellation policy<br>
H1: The cancellation policy has no effect on the average booking rate<br>
<br>*Assumption: The monthly booking rate here is based on reviews per month. This implies that its not actually the booking rate(including cancellation) but its the booking rate where people have actually stayed at the accomodation(or the staying rate)* <br>

We can find this out using a One-Way Anova Test.We used variable reviews_per_month as a measure for the bookings rate.<br>

```{r,warning=FALSE,message=FALSE}
# Visualize the distribution of reviews_per_month for different cancellation policy using a boxplot
library(ggpubr)
ggboxplot(Chicago, x = "cancellation_policy", y = "reviews_per_month", 
          color = "cancellation_policy",
          ylab = "Bookings per month", xlab = "Cancellation Policy")

```

<br>Initial inspection of the data using the boxplot suggests that there are differences in the Booking rate for different cancellation policies but it is not very clear so that we are not able to determine whether all are different or only a few are different. To investigate these differences we then fited an one-way ANOVA model.<br>

```{r,set-options}
options(width = 90)
# Compute the analysis of variance
anova1 <- aov(reviews_per_month ~ cancellation_policy, data = Chicago)

# Summary of the analysis
summary(anova1)
```

As the p-value is less than the significance level 0.05 in the model summary, we can conclude that there are significant differences among the groups of cancellation policy.<br>

In one-way ANOVA test, a significant p-value indicates that some of the group means are different, but we don't know which pairs of groups are different. <br>

As the ANOVA test is significant, we can compute Tukey HSD for performing multiple pairwise-comparisons between the means of pairs of groups.

```{r}
TukeyHSD(anova1)
```

<br>**We can see from the results that the average booking rate is significantly different for moderate and flexible cancellation policies as well as Strict and flexible cancellation policies.** <br>

**Therefore, our null hypothesis that the booking rate changes with the cancellation policy holds true. As we can see from the box plot, the accomodations with Strict and Moderate cancellation policies have higher average booking rate(staying rate)**<br>

###Analysis 2: Comparing the Review Scores based on bed type

<br>This analysis is done to see if the type of bed provided by a host in AirBNB has any effects on the average review score for that listing<br>

H0: The average review score changes with different bed types<br>
H1: There is no effect of bed type on average review scores<br>

<br>*Assumption: Different bed types are assumed to be providing different comfort levels(bed types are proportional to comfort)* <br>


```{r}
# Visualizing the distribution of review_scores_rating for different bed type using a boxplot
ggboxplot(Chicago, x = "bed_type", y = "review_scores_rating", 
          color = "bed_type",
          ylab = "Review Scores Rating", xlab = "Bed Type")
```

<br>Initial inspection of the data using the boxplot suggests that there are no differences in the review rating for different bed types but it is not clear enough to determine whether all are similar or only a few are similar. To investigate these differences we fit the one-way ANOVA model.<br>

```{r}
# Compute the analysis of variance
anova2 <- aov(review_scores_rating ~ bed_type, data = Chicago)

# Summary of the analysis
summary(anova2)
```

<br>**Therefore, we reject the null hypothesis that the average review scores changes based on the bed types. That's not quite we epxpected based on our intuition. Intuitively, people prefer comfort and hence would prefer "Real Bed" versus lets say "Sofa" or "Couch".**<br>
<br>**However our results imply that when it comes to high average review scores, there is no significant effect of bed types on it. Also this shows that people tends to base their ratings not just on the comfort**<br>


###Analysis 3: Predicting the prices of listings & Finding factors impacting prices

```{r,echo=FALSE}
options(width = 100)
##lm
price_model <- lm(price ~ as.factor(neighbourhood_cleansed) + host_since + Breakfast + Kitchen + Pets + AirCondition + Internet + TV + reviews_per_month + review_scores_rating + number_of_reviews + review_scores_accuracy + review_scores_value + review_scores_cleanliness + review_scores_checkin + review_scores_communication + review_scores_location + as.factor(cancellation_policy) + guests_included + beds + as.factor(bed_type) + bathrooms + as.factor(property_type) + as.factor(room_type), 
   data = Chicago)

summary(price_model)
```

Following are the results that we attain from the Linear Reression model along with the hypothesis/explaination of the behavior of the variables:

<br>**Based on the p-values for the above regression model, we can say that amenities like TV and Air Conditioning significantly impact the prices. Since the Beta-values are positive, we can say that TV and AC leads to increase in prices. This can be because AC and TV are comparatively  expensive ammenties**<br>

<br>**Also, the booking rate is inversely proportional to the prices considering the other factors remains unchanged.**<br>

<br>**Number of guests, number of beds and number of bathrooms directly effect the prices. Generally in the hotels as well, as the number of guests and beds increases, prices increases(considering amenities/comfort and location is same)**<br>

<br>**The categorical variables : neighbourhood, property type and room type have significant impact on prices. This is generally true as properties in posh localities will be more expensive where as the similar listings away from the expensive housing areas will be cheaper. Also, a shared room will be cheaper than whole house/apartment in the same locality.**<br>
